\bf ISODS: An ISODATA-Based Initial Centroid Algorithm

### 2021-01-08 paper/paper.tex
add experiment 2 results and conclusion

\documentclass[a4paper,10pt]{article}
\usepackage{graphicx}
\begin{document}

@@

@@

\begin{table}
      \begin{tabular}{ |p{1.25cm}||p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm}|}
            \hline
            \multicolumn{7}{|c|}{\textbf{ISODATA}}                               \\
            \hline
            Init. & Iter.  & Clusters & ARI             & RI    & MI    & HI     \\
            \hline
            \multicolumn{7}{|c|}{\textit{Iris}}                                  \\[1ex]
            ISODS & 13.46  & 2.599    & \textbf{0.646}  & 0.826 & 0.173 & 0.653  \\
            KM++  & 15.43  & 2.544    & 0.628           & 0.816 & 0.184 & 0.631  \\
            Rand  & 11.90  & 2.41     & 0.607           & 0.802 & 0.198 & 0.604  \\
            \hline
            \multicolumn{7}{|c|}{\textit{Glass (Normalized)}}                    \\[1ex]
            ISODS & 3.19   & 1.066    & \textbf{0.0018} & 0.266 & 0.734 & -0.468 \\
            KM++  & 4.02   & 1.028    & 0.0011          & 0.265 & 0.735 & -0.470 \\
            Rand  & 3.69   & 1.055    & 0.0013          & 0.266 & 0.733 & -0.467 \\
            \hline
            \multicolumn{7}{|c|}{\textit{Wine}}                                  \\[1ex]
            ISODS & 9.16   & 1.836    & \textbf{0.313}  & 0.618 & 0.382 & 0.236  \\
            KM++  & 8.76   & 1.816    & 0.307           & 0.612 & 0.388 & 0.224  \\
            Rand  & 9.26   & 1.810    & 0.304           & 0.608 & 0.391 & 0.217  \\
            \hline 
            \multicolumn{7}{|c|}{\textbf{K-Means}}                               \\
            \hline
            \multicolumn{7}{|c|}{\textit{Iris}}                                  \\[1ex]
            ISODS & 2.29   & 3        & \textbf{0.623}  & 0.833 & 0.167 & 0.667  \\
            KM++  & 4.149  & 3        & 0.584           & 0.811 & 0.189 & 0.621  \\
            Rand  & 29.402 & 3        & 0.569           & 0.797 & 0.203 & 0.593  \\
            \hline
            \multicolumn{7}{|c|}{\textit{Glass (Normalized)}}                    \\[1ex]
            ISODS & 5.651  & 6        & 0.164           & 0.665 & 0.335 & 0.329  \\
            KM++  & 4.866  & 6        & \textbf{0.173}  & 0.643 & 0.357 & 0.285  \\
            Rand  & 77.178 & 6        & 0.151           & 0.634 & 0.366 & 0.268  \\
            \hline
            \multicolumn{7}{|c|}{\textit{Wine}}                                  \\[1ex]
            ISODS & 3.145  & 3        & \textbf{0.882}  & 0.947 & 0.053 & 0.895  \\
            KM++  & 5.388  & 3        & 0.839           & 0.926 & 0.074 & 0.851  \\
            Rand  & 24.775 & 3        & 0.834           & 0.920 & 0.080 & 0.840  \\
            \hline
      \end{tabular}
      \caption{UCI benchmark dataset results averaged over 1000 iterations of
            ISODATA and K-Means comparing three different initial cluster center
            algorithms (\textit{Init.}). Results in the table contain
            information on average iterations required (\textit{Iter.}), average
            number of clusters created (\textit{Clusters}), and validity
            measures.}
      \label{tab:ucitable}
\end{table}

@@

@@

\section{Conclusion}
Put Conclusion here.

@@

\begin{center}
      {\large \bf References}
\end{center}

>>>>

\documentclass[a4paper,10pt]{article}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\graphicspath{ {./img/} }
\begin{document}

@@

\begin{table}[ht]
      \begin{tabular}{ |p{1.25cm}||p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm}|}
            \hline
            \multicolumn{7}{|c|}{\textbf{ISODATA}}                               \\
            \hline
            Init. & Iter.  & Clusters & ARI             & RI    & MI    & HI     \\
            \hline
            \multicolumn{7}{|c|}{\textit{Iris}}                                  \\[1ex]
            ISODS & 13.46  & 2.599    & \textbf{0.646}  & 0.826 & 0.173 & 0.653  \\
            KM++  & 15.43  & 2.544    & 0.628           & 0.816 & 0.184 & 0.631  \\
            Rand  & 11.90  & 2.41     & 0.607           & 0.802 & 0.198 & 0.604  \\
            \hline
            \multicolumn{7}{|c|}{\textit{Glass (Normalized)}}                    \\[1ex]
            ISODS & 3.19   & 1.066    & \textbf{0.0018} & 0.266 & 0.734 & -0.468 \\
            KM++  & 4.02   & 1.028    & 0.0011          & 0.265 & 0.735 & -0.470 \\
            Rand  & 3.69   & 1.055    & 0.0013          & 0.266 & 0.733 & -0.467 \\
            \hline
            \multicolumn{7}{|c|}{\textit{Wine}}                                  \\[1ex]
            ISODS & 9.16   & 1.836    & \textbf{0.313}  & 0.618 & 0.382 & 0.236  \\
            KM++  & 8.76   & 1.816    & 0.307           & 0.612 & 0.388 & 0.224  \\
            Rand  & 9.26   & 1.810    & 0.304           & 0.608 & 0.391 & 0.217  \\
            \hline 
            \multicolumn{7}{|c|}{\textbf{K-Means}}                               \\
            \hline
            \multicolumn{7}{|c|}{\textit{Iris}}                                  \\[1ex]
            ISODS & 2.29   & 3        & \textbf{0.623}  & 0.833 & 0.167 & 0.667  \\
            KM++  & 4.149  & 3        & 0.584           & 0.811 & 0.189 & 0.621  \\
            Rand  & 29.402 & 3        & 0.569           & 0.797 & 0.203 & 0.593  \\
            \hline
            \multicolumn{7}{|c|}{\textit{Glass (Normalized)}}                    \\[1ex]
            ISODS & 5.651  & 6        & 0.164           & 0.665 & 0.335 & 0.329  \\
            KM++  & 4.866  & 6        & \textbf{0.173}  & 0.643 & 0.357 & 0.285  \\
            Rand  & 77.178 & 6        & 0.151           & 0.634 & 0.366 & 0.268  \\
            \hline
            \multicolumn{7}{|c|}{\textit{Wine}}                                  \\[1ex]
            ISODS & 3.145  & 3        & \textbf{0.882}  & 0.947 & 0.053 & 0.895  \\
            KM++  & 5.388  & 3        & 0.839           & 0.926 & 0.074 & 0.851  \\
            Rand  & 24.775 & 3        & 0.834           & 0.920 & 0.080 & 0.840  \\
            \hline
      \end{tabular}
      \caption{UCI benchmark dataset results averaged over 1000 iterations of
            ISODATA and K-Means comparing three different initial cluster center
            algorithms (\textit{Init.}). Results in the table contain
            information on average iterations required (\textit{Iter.}), average
            number of clusters created (\textit{Clusters}), and validity
            measures.}
      \label{tab:ucitable}
\end{table}

@@

@@

\begin{figure}[ht]
      \centering
      \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\linewidth]{salinasgt.png}
            \caption{Salinas-A Ground Truth}
            \label{fig:salinasgt}
      \end{subfigure}
      \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\linewidth]{risodata.png}
            \caption{Random Centers (ARI=0.797)}
            \label{fig:salinasr}
      \end{subfigure}\hfill
      \centering
      \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\linewidth]{kisodata.png}
            \caption{KM++ Centers (ARI=0.781)}
            \label{fig:salinask}
      \end{subfigure}
      \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\linewidth]{iisodata.png}
            \caption{ISODS Centers (ARI=0.758)}
            \label{fig:salinasi}
      \end{subfigure}\hfill \caption{Heatmap showing cluster assignments of each
            pixel using ISODATA with different initial cluster center methods.}
      \label{fig:salinas}
\end{figure}

Following the results on the benchmark datasets, we applied ISODATA to the
Salinas-A hyperspectral image dataset \cite{Salinas}. Salinas-A is a small
subscene of the Salinas image. It has a size of $86$ pixels by $83$ pixels, each
pixel contains $224$ bands, and each pixel is assigned to $1$ of $6$ classes,
disregarding the background. Due to the large number of dimensions in this
dataset, prinicipal component analysis (PCA) is used to extract features
\cite{Rodarmel2002}. In this experimental setup, we use $3$ prinicipal
components as the input to the model. We utilize the same 80/20 test-train split
as the previous experiment and Z-Score Normalize the data. Additionally, the
same ISODATA parameter selection approach is employed. Then, we run ISODATA for
$1000$ iterations using ISODS, K-Means++ Centers, and Random Centers. We save
the best performing model (based on test data performance) for each initial
cluster center approach and fit each pixel to the original dataset to obtain the
pictures shown in Figure \ref{fig:salinas}. Finally, the external cluster
validation measures are shown in Table \ref{tab:salinastable}.

\begin{table}[ht]
      \begin{tabular}{ |p{1.25cm}||p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm}|}
            \hline
            \multicolumn{7}{|c|}{\textbf{ISODATA} - \textit{Salinas-A}}        \\
            \hline
            Init. & Iter. & Clusters & ARI            & RI    & MI    & HI     \\
            \hline
            \multicolumn{7}{|c|}{\textit{Average}}                             \\[1ex]
            ISODS & 2.772 & 2.42     & \textbf{0.156} & 0.388 & 0.612 & -0.224 \\
            KM++  & 3.177 & 1.097    & 0.019          & 0.228 & 0.772 & -0.544 \\
            Rand  & 3.693 & 1.584    & 0.084          & 0.308 & 0.691 & -0.383 \\
            \hline
            \multicolumn{7}{|c|}{\textit{Best}}                                \\[1ex]
            ISODS & 2     & 6        & 0.758          & 0.921 & 0.078 & 0.842  \\
            KM++  & 8     & 6        & 0.781          & 0.930 & 0.070 & 0.860  \\
            Rand  & 14    & 6        & \textbf{0.798} & 0.933 & 0.066 & 0.867  \\
            \hline
      \end{tabular}
      \caption{Salinas-A dataset average and best results over 1000 iterations
            of ISODATA comparing three different initial cluster center
            algorithms (\textit{Init.}). Results in the table contain
            information on iterations required (\textit{Iter.}), number of
            clusters created (\textit{Clusters}), and validity measures.}
      \label{tab:salinastable}
\end{table}

@@

\section{Conclusion}
This study presents a method for generating initial cluster centers based on a
splitting approach, ISODS. We tested ISODS on bechmark datasets as well as the
Salinas-A hyperspectral image dataset. Both tests on average shows the
performance of ISODS using external cluster validation measures, especially on
the \textit{adjusted rand index} measure. Additonally, ISODS may decrease the
number of iterations required for a clustering algorithm such as ISODATA or
K-Means to converge. However, due to the complexity of the ISODS approach,
specifically calculating the means of the clusters, the number of iterations
taken by ISODS should be considered. When used in conjunction with ISODATA,
ISODS also shows an average number of clusters generated closer to the expected
output.

@@
